Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de~Rham complex over convex domains and domains with shellable triangulations

### 2025-01-15 main.tex
further TODOs done

The qualitative existence of Poincar\'e--Friedrichs inequalities for the differential operators of vector calculus and exterior calculus is known for a large class of domains. 
There is a particularly large body of literature on the Poincar\'e inequality for the gradient and its numerous variants,
which may include weighted integrals or boundary terms. 
These Poincar\'e-type inequalities are often derived via non-constructive arguments, such as the Rellich embedding theorem, 
% TODO {[TCF: I believe this is often refered to as the Tartare-Petree theorem]}
and thus the resulting constants are not explicitly computable.

@@

% veeser and verfurth 
% sequential methodology
There are numerous estimates for Poincar\'e--Friedrichs constants that only rely on locally computable geometric quantities, such as the diameter and volumes of simplices. 
Veeser and Verf\"urth~\cite{veeser2012poincare} provide computable upper bounds in the case of the classical Sobolev space $W^{1,p}(\Omega)$ over triangulated domains, with focus on efficient estimates for vertex stars. Naturally, their estimates depend on the shape regularity of the mesh.
A whole class of upper bounds for Poincar\'e--Friedrichs inequalities uses some form of passing through the triangulation and constructs the potentials step-by-step. 
The underlying idea is that we first construct a potential for the gradient over an initial simplex. 
Every time we have found a potential over a subdomain, we construct a potential over a neighboring simplex or patch:
along the interfacing intersection, the two potentials will only differ by a constant, 
and that difference can easily be removed to ensure continuity across that interface. 
Cell by cell, the potential is constructed over increasing subdomains, matching along the interfacing intersections, until the entire domain is exhausted.  
The method is known in the finite element literature~\cite{ern2021finite}. 
% TODO {[TCF: Could you reference a particular theorem or chapter?]} 
It was previously used in the context of finite volume methods~\cite{Eym_Gal_Her_00}, broken (weakly continuous) Sobolev spaces~\cite{vohralik2005discrete}, or more recently in continuous--discrete comparison results~\cite{Brae_Pill_Sch_p_rob_09, ern2020stable, Chaum_Voh_p_rob_3D_H_curl_24,  Voh_loc_glob_H1_24}. 
This sequential procedure applies to general triangulated domains, not only local stars, though the latter are our main interest here. 
Most importantly, these sequential estimates of Poincar\'e--Friedrichs constants generally circumvent the effect of low boundary regularity and only rely on the shape regularity.

@@

% TODO: Figure out whether to use the following paragraph.
% Our main objective remains to find potentials for the differential operators of vector calculus and exterior calculus, such as the curl and divergence, over triangulated domains. 
% The operator norms of these potentials will serve as computable Poincar\'e--Friedrichs constants. 
% In light of the different approaches listed before, 
% we seek constants that can be explicitly computed in terms of the mesh geometry, that do not require solving global finite element problems, and that are independent of the boundary regularity of the domain.

@@

% TODO: check whether introduction is readable
% In the gradient case of Section~\ref{subsection:intro_grad}, it is possible to perform a simple pass through the triangulation, 
% where merely each subsequent simplex has to be connected to the previous ones by a $(n-1)$-dimensional face. 
% We always construct a gradient potential over the new simplex, and since the gradient kernel is trivial, 
% formed merely by constants, it is enough to fix this constant using the value already known on the $(n-1)$-dimensional face.

@@

    % TODO: Change colors to grey scale

@@

    \draw[line width=0.1cm] (O) -- (A) -- (E) -- cycle;
  \end{tikzpicture}
\end{tabular} 
\caption{From left to right: local patches around a vertex, and edge, and a triangle. The local patch of a full-dimensional simplex is the only the simplex itself (and its subsimplices).}\label{figure:patches}
\end{figure}

@@

\begin{remark}
    The ratio $\aspectratio(T)$ measures the ``shape quality'' of an $n$-dimensional simplex $T$ and is an instance of a so-called \emph{shape measure}.
    For example, the reference triangle has aspect shape measure $2$ and the reference tetrahedron has aspect shape measure $\sqrt{6}$. 
    Numerous alternative shape measures have been used throughout the literature of numerical analysis and computational geometry to quantify the quality of simplices (see~\cite[p.61, Definition 5.1]{braess2001finite}, 
    \cite[p.97, Definition 4.2.16 ]{brenner2008mathematical}, % TODO: add brackets here.
    \cite[Definition~11.2]{ern2021finite}). 
\end{remark}

We gather a few relationships between geometric and algebraic entities and compare the different shape measures of a single simplex.
% DONE {Settle the subscripts of the vector norm.}
% TODO {Replace the constant Ceins and Czwei by the square root of n immediately.}

@@

% We discuss the Sobolev spaces of vector calculus and exterior calculus 
% with particular emphasis on their transformation behavior. 
% We refer the reader to Ern and Guermond~\cite{ern2021finite} and Hiptmair~\cite{hiptmair2002finite} for further background material on Sobolev vector analysis. 
% Exterior algebra and exterior products are reviewed in the textbooks by Greub~\cite{greub1967multilinear} and Lee~\cite{lee2012smooth}.
% TODO: Fuse these two paragraphs

We review in this section the Sobolev spaces of vector and exterior calculus and their transformation behavior. We refer the reader to Ern and Guermond~\cite{ern2021finite} and Hiptmair~\cite{hiptmair2002finite} for background material on Sobolev vector analysis and to Greub~\cite{greub1967multilinear} and Lee~\cite{lee2012smooth} for exterior algebra and exterior products.

@@

% \begin{lemma} % TODO: possibly useful
%     Let $1 \leq p < \infty$. A sequence converges to $u \in L^{p}$ if and only if it converges locally by measure and the $L^{p}$ norms converge.
% \end{lemma}

@@

\begin{lemma}\label{lemma:connectivity}
    Let $\calT$ be a finite $n$-dimensional simplicial complex whose underlying space is a manifold $\calM$.
    If the underlying space of $\calT$ is connected, then $\calT$ is face-connected.
\end{lemma}
\begin{proof}
    We first show that each vertex star is face-connected via a short induction argument.
    Clearly, any simplicial $1$-ball and simplicial $1$-sphere are face-connected. 
    Now, if $n \geq 1$ and $V \in \calT$,
    then the $n$-simplices in $\patch_{\calT}(V)$ are in correspondence to the $(n-1)$-simplices in $\carapace_{\calT}(V)$.
    Hence $\patch_{\calT}(V)$, a triangulation of dimension $n-1$, is face-connected if and only if $\carapace_{\calT}(V)$, a triangulation of dimension $n-1$, is face-connected.
    % then any vertex star in any $(n+1)$-dimensional manifold triangulation is already face-connected if all simplicial $n$-balls and $n$-spheres are face-connected.
    The induction argument implies that each vertex star in $\calT$ is face-connected.

@@

\begin{lemma}\label{lemma:shell_triang_man}
    Let $\calT$ be an $n$-dimensional simplicial complex 
    with a shelling $T_{0}, T_{1}, T_{2}, \dots, T_{M}$,
    such that each simplex of dimension $n-1$ is contained in at most two simplices. 
    % DONE: what is non-branching?
    Then
    $$
        X_{m} := T_{0} \cup T_{1} \cup \dots \cup T_{m}
    $$ 
    is a triangulated manifold with boundary for all $0 \leq m \leq M$.
    % 
    In particular, $X_{m}$ is a topological $n$-ball when $m < M$, 
    and 
    $X_{M}$ is either a topological $n$-ball or topological $n$-sphere. 
\end{lemma}
\begin{proof}  
    We prove this claim by induction. 
    Certainly, if $\calT$ contains only one single $n$-simplex, then it is a shellable triangulation of a topological $n$-ball. Let $1 \leq m \leq M$ and suppose that 
    \begin{align*}
        X_{m-1} := T_{0} \cup T_{1} \cup \dots \cup T_{m-1}
    \end{align*}
    is a topological $n$-ball. Let $T_{m}$ is the next $n$-simplex in the shelling.
    By definition, $\Gamma_{m} := X_{m-1} \cap T_{m}$ is a submanifold of $\partial T_{m}$,
    and it is triangulated by some faces of $T_{m}$ and their subsimplices.

@@

    On the one hand, 
    if $\Gamma_{m}$ is the entire boundary of $T_{m}$, 
    then it must also be the boundary of the topological $n$-ball $X_{m-1}$. 
    Hence, $X_{m-1} \cup T_{m}$ must be a topological $n$-sphere and thus a manifold without boundary, 
    which requires $m = M$.
    % TODO {TCF: I don't understand why XM couldn't be a topological n-ball.}
    On the other hand, 
    if $\Gamma_{m}$ is a proper subset of the entire boundary of $T_{m}$, 
    then $X_{m-1} \cup T_{m}$ is still a topological $n$-ball.
\end{proof}

@@

        The vector $a$ is the difference between $\hat h_{G}$ and $\hat h_{K}$ after both are rescaled such that they lie on the affine hyperplane with normal vector $\hat x$. That difference has length at most $\tan(\alpha_{G}) + \tan(\alpha_{K})$.
        We calculate 
        \begin{align*}
            \vecnorm{a}
            = 
            \left\|
            \frac{\vecnorm{ x_{G} }}{\vecnorm{ h_{G} }} \hat h_{{G}} 
            - 
            \frac{\vecnorm{ x_{K} }}{\vecnorm{ h_{K} }} \hat h_K
            \right\|
            &
            \leq 
            \frac{\vecnorm{ x_{G} }}{\vecnorm{ h_{G} }} 
            + 
            \frac{\vecnorm{ x_{K} }}{\vecnorm{ h_{K} }} 
            % \\&
            \leq 
            (\ell+1) 
            \left( 1 + \mu_{T,\ell} \right)
            \sup_{ K \in \patch_{\calT}(S) }
            \aspectratio(K) % \frac{ \diam(K)^{n} }{ n! \vol(K) }
            .
        \end{align*}
        We define 
        \begin{align}
            \nu_{T,\ell} := \sqrt{ 1 + \frac{ (\ell+1)^{2}\left( 1 + \mu_{T,\ell} \right)^{2} \aspectratio(\calT)^{2} }{4} } + \frac{ (\ell+1)\left( 1 + \mu_{T,\ell} \right) \aspectratio(\calT) }{2}
            .
        \end{align}
        In summary, 
        the singular values of the Jacobian of $\Psi$ at almost every $x$ satisfy 
        % using that $\sqrt{s^2+t^2} \leq s + t$ for any $s,t \geq 0$, 
        \begin{align}
            \sigma_{1}(\Psi,x) % \matnorm{ \Jacobian\Psi(x) }
            &\leq  
            (\ell+1)
            \diameterratio(\calT)
            \aspectratio(T)\mu_{T,\ell} % \frac{ \diam(T)^{n} }{ n! \vol(T) }
            \cdot 
            \nu_{T,\ell}
            ,
            \\
            \sigma_{n}(\Psi,x)^{-1} % \matnorm{ \Jacobian\Psi(z)^{-1} }
            &\leq  
            (\ell+1)
            \diameterratio(\calT)
            \aspectratio(\calT) 
            \cdot 
            \nu_{T,\ell}
            ,
            \\
            \sigma_{  2}(\Psi,x) 
            &= 
            \dots
            =
            \sigma_{n-1}(\Psi,x) 
            \leq 
            (\ell+1) \diameterratio(\calT) \aspectratio(T)\mu_{T,\ell}
            ,
            \\
            \sigma_{n-1}(\Psi,x)^{-1} 
            &= 
            \dots
            =
            \sigma_{  2}(\Psi,x)^{-1} 
            \leq 
            (\ell+1) \diameterratio(\calT) \aspectratio(T) 
            ,
            \\
            \det\Jacobian\Psi(x)
            &\leq 
            \left( 
            (\ell+1)
            \diameterratio(\calT)
            \aspectratio(T)\mu_{T,\ell} % \frac{ \diam(T)^{n} }{ n! \vol(T) }
            \right)^{n},
            \\
            \det\Jacobian\Psi^{-1}(x)
            &\leq 
            \left( 
            (\ell+1)
            \diameterratio(\calT)
            \aspectratio(\calT) 
            \right)^{n}
            .
        \end{align}
        We have shown that $\Psi$ and $\Psi^{-1}$ are piecewise Lipschitz with respect to some essentially disjoint closed partitions over their respective domains.
        % It remains to show both are Lipschitz, which we illustrate for $\Psi$, the argument for $\Psi^{-1}$ being analogous. 
        % Rademacher's theorem then shows that both have essentially bounded weak Jacobians. 
        Every function that is piecewise Lipschitz with respect to the triangulation $\calK$ must be Lipschitz over all of $\underlying{\calK}$, since there exists a bi-Lipschitz transformation from $\underlying{\calK}$ onto the unit ball.
        % We use a technique similar to Lemma~\ref{lemma:stars_are_shellable}. 
        % After application of a piecewise linear Lipschitz mapping, 
        % we can assume that $\underlying{\calK}$ is convex without loss of generality. 
        % Any function piecewise Lipschitz with respect to $\calK$ must be Lipschitz over $\underlying{\calK}$. 
        % \mwl{TODO: not enough. Instead, use the homeomorphism onto a ball and estimate its Lipschitz constant.}

@@

        The singular values $\sigma_{\max}(\Theta,G) \geq 1$ and $\sigma_{\min}(\Theta,G) \leq 1$ is monotonically increasing and decreasing, respectively, in $\tan(\beta)$. Hence they must the extremal maximal singular values of $\Jacobian\Theta_{|G}$.

@@

        Recall that $G \in \calG$ is obtained from $T$ via barycentric subdivisions: 
        first at $z_{S}$, leading to some intermediate $n$-simplex $T'$, and then at $z_{S'}$, leading to $G \subseteq T'$. 
        Let $F_{z} \subseteq G$ be the face opposite to the vertex $z_{S'}$, which is some face of $T'$.
        Now, the height of $F_{z}$ in $G$ is $(n-\ell)$-th of the height of $F_{z}$ in $T'$
        since $S'$ has dimension $n-\ell-1$. 
        The height of $F_{z}$ in $T'$ is just the height of the its opposing vertex, which lies in $S'$,
        and which equals the height of that vertex in $T$.
%         which must be contained in some face of $T$ containing $S$. 
%         The height vector $h_{z}$ is orthogonal to $F_{z}$. 
%         Since $S'$ has dimension $n-\ell-1$, the vector $(n-\ell) h_{z}$ is the height vector of that face of $T$,
%         pointing to some vertex of $S'$ and standing orthogonally on the face opposite to that vertex. 
        Thus, 
        \begin{align*}
            \frac{ \vecnorm{ b_{z} } }{ \vecnorm{ h_{z} } }
            = 
            \frac{ \vecnorm{ b_{z} } }{ (n-\ell)^{-1} \vecnorm{ (n-\ell) h_{z} } }
            = 
            \frac{ \vecnorm{ z_{S'} } }{ (n-\ell)^{-1} \vecnorm{ (n-\ell) h_{z} } }
            \leq 
            (n-\ell) \aspectratio(T)
%             = 
%             \frac{ \vecnorm{ b_{z} } \vol(F_{z}) }{ (n-\ell)^{-1} n \vol(T) }
%             \leq 
%             \frac{ n-\ell }{n!}
%             \cdot 
%             \frac{ \diam(T)^{n} }{ \vol(T) }
            .
        \end{align*}
        We abbreviate 
        \begin{align}\label{math:upper_bound_for_theta}
            \mu_{T,\ell} 
            := 
            \frac 1 2 \sqrt{ \left( 1 + \rho \right)^{2} + (1+\rho)^{2} (n-\ell)^{2} \aspectratio(T)^{2} } 
            + 
            \frac 1 2 \sqrt{ \left( 1 - \rho \right)^{2} + (1+\rho)^{2} (n-\ell)^{2} \aspectratio(T)^{2} }
            .
        \end{align}
        This establishes bounds on the singular values of the transformation. 
        In summary, the singular values of the Jacobian of $\Theta$ at almost every $x$ satisfy 
        \begin{align}
            \sigma_{1}(\Theta,x) \leq \mu_{T,\ell},
            \quad 
            \sigma_{2}(\Theta,x) = \dots = \sigma_{n-1}(\Theta) = 1,
            \quad 
            \sigma_{n}(\Theta,x)^{-1} = \frac{ \sigma_{1}(\Theta,x) }{ \left| \det \Jacobian\Theta_{|G} \right| } \leq \mu_{T,\ell} / \rho.
        \end{align}

>>>>

The qualitative existence of Poincar\'e--Friedrichs inequalities for the differential operators of vector calculus and exterior calculus is known for a large class of domains. 
There is a particularly large body of literature on the Poincar\'e inequality for the gradient and its numerous variants,
which may include weighted integrals or boundary terms. 
These Poincar\'e-type inequalities are often derived via non-constructive arguments, such as the Rellich embedding theorem, 
% DONT {[TCF: I believe this is often refered to as the Tartare-Petree theorem]}
and thus the resulting constants are not explicitly computable.

@@

% veeser and verfurth 
% sequential methodology
There are numerous estimates for Poincar\'e--Friedrichs constants that only rely on locally computable geometric quantities, such as the diameter and volumes of simplices. 
Veeser and Verf\"urth~\cite{veeser2012poincare} provide computable upper bounds in the case of the classical Sobolev space $W^{1,p}(\Omega)$ over triangulated domains, with focus on efficient estimates for vertex stars. Naturally, their estimates depend on the shape regularity of the mesh.
A whole class of upper bounds for Poincar\'e--Friedrichs inequalities uses some form of passing through the triangulation and constructs the potentials step-by-step. 
The underlying idea is that we first construct a potential for the gradient over an initial simplex. 
Every time we have found a potential over a subdomain, we construct a potential over a neighboring simplex or patch:
along the interfacing intersection, the two potentials will only differ by a constant, 
and that difference can easily be removed to ensure continuity across that interface. 
Cell by cell, the potential is constructed over increasing subdomains, matching along the interfacing intersections, until the entire domain is exhausted.  
The method is known in the finite element literature~\cite{ern2021finite}. 
% DONE {[TCF: Could you reference a particular theorem or chapter?]} 
It was previously used in the context of finite volume methods~\cite{Eym_Gal_Her_00}, broken (weakly continuous) Sobolev spaces~\cite{vohralik2005discrete}, or more recently in continuous--discrete comparison results~\cite{Brae_Pill_Sch_p_rob_09, ern2020stable, Chaum_Voh_p_rob_3D_H_curl_24,  Voh_loc_glob_H1_24}. 
This sequential procedure applies to general triangulated domains, not only local stars, though the latter are our main interest here. 
Most importantly, these sequential estimates of Poincar\'e--Friedrichs constants generally circumvent the effect of low boundary regularity and only rely on the shape regularity.

@@

% DONE: Figure out whether to use the following paragraph.
% Our main objective remains to find potentials for the differential operators of vector calculus and exterior calculus, such as the curl and divergence, over triangulated domains. 
% The operator norms of these potentials will serve as computable Poincar\'e--Friedrichs constants. 
% In light of the different approaches listed before, 
% we seek constants that can be explicitly computed in terms of the mesh geometry, that do not require solving global finite element problems, and that are independent of the boundary regularity of the domain.

@@

% DONE: check whether introduction is readable
% In the gradient case of Section~\ref{subsection:intro_grad}, it is possible to perform a simple pass through the triangulation, 
% where merely each subsequent simplex has to be connected to the previous ones by a $(n-1)$-dimensional face. 
% We always construct a gradient potential over the new simplex, and since the gradient kernel is trivial, 
% formed merely by constants, it is enough to fix this constant using the value already known on the $(n-1)$-dimensional face.

@@

@@

    \draw[line width=0.1cm] (O) -- (A) -- (E) -- cycle;
  \end{tikzpicture}
\end{tabular} 
\caption{From left to right: local patches around a vertex, and edge, and a triangle. The local patch of any full-dimensional simplex only consists of that simplex itself (and its subsimplices).}\label{figure:patches}
\end{figure}

@@

\begin{remark}
    The ratio $\aspectratio(T)$ measures the ``shape quality'' of an $n$-dimensional simplex $T$ and is an instance of a so-called \emph{shape measure}.
    For example, the reference triangle has aspect shape measure $2$ and the reference tetrahedron has aspect shape measure $\sqrt{6}$. 
    Numerous alternative shape measures have been used throughout the literature of numerical analysis and computational geometry to quantify the quality of simplices (see~\cite[p.61, Definition 5.1]{braess2001finite}, 
    {\cite[p.97, Definition 4.2.16 ]{brenner2008mathematical}, % TODO: add brackets here in the final version or figure out how to fix the display in the editor.
    \cite[Definition~11.2]{ern2021finite}). 
\end{remark}

We gather a few relationships between geometric and algebraic entities and compare the different shape measures of a single simplex.
% DONE {Settle the subscripts of the vector norm.}
% DONT {Replace the constant Ceins and Czwei by the square root of n immediately.}

@@

We review in this section the Sobolev spaces of vector and exterior calculus with particular emphasis on their transformation behavior. 
We refer the reader to Ern and Guermond~\cite{ern2021finite} and Hiptmair~\cite{hiptmair2002finite} for background material on Sobolev vector analysis and to Greub~\cite{greub1967multilinear} and Lee~\cite{lee2012smooth} for exterior algebra and exterior products.

@@

% \begin{lemma} % DONT: possibly useful
%     Let $1 \leq p < \infty$. A sequence converges to $u \in L^{p}$ if and only if it converges locally by measure and the $L^{p}$ norms converge.
% \end{lemma}

@@

\begin{lemma}\label{lemma:connectivity}
    Let $\calT$ be a finite $n$-dimensional simplicial complex whose underlying space is a manifold $\calM$.
    If the underlying space of $\calT$ is connected, then $\calT$ is face-connected.
\end{lemma}
\begin{proof}
    We first show that each vertex star is face-connected via a short induction argument.
    Clearly, any simplicial $1$-ball and simplicial $1$-sphere are face-connected. 
    Now, if $n \geq 1$ and $V \in \calT$,
    then the $n$-simplices in $\patch_{\calT}(V)$ are in correspondence to the $(n-1)$-simplices in $\carapace_{\calT}(V)$.
    Hence, $\patch_{\calT}(V)$, a triangulation of dimension $n-1$, is face-connected if and only if $\carapace_{\calT}(V)$, a triangulation of dimension $n-1$, is face-connected.
    % then any vertex star in any $(n+1)$-dimensional manifold triangulation is already face-connected if all simplicial $n$-balls and $n$-spheres are face-connected.
    The induction argument implies that each vertex star in $\calT$ is face-connected.

@@

\begin{lemma}\label{lemma:shell_triang_man}
    Let $\calT$ be an $n$-dimensional simplicial complex 
    with a shelling $T_{0}, T_{1}, T_{2}, \dots, T_{M}$,
    such that each simplex of dimension $n-1$ is contained in at most two simplices. 
    % DONE: what is non-branching?
    Then
    $$
        X_{m} := T_{0} \cup T_{1} \cup \dots \cup T_{m}
    $$ 
    is a triangulated manifold with boundary for all $0 \leq m \leq M$.
    % 
    In particular, $X_{m}$ is a topological $n$-ball when $m < M$, 
    and 
    $X_{M}$ is either a topological $n$-ball or topological $n$-sphere. 
\end{lemma}
\begin{proof}  
    We prove this claim by induction. 
    Certainly, if $\calT$ contains only one single $n$-simplex, then it is a shellable triangulation of a topological $n$-ball. 
    Next, let $1 \leq m \leq M$ and suppose that
    \begin{align*}
        X_{m-1} := T_{0} \cup T_{1} \cup \dots \cup T_{m-1}
    \end{align*}
    is already known to be a topological $n$-ball. Let $T_{m}$ be the next $n$-simplex in the shelling.
    By definition, $\Gamma_{m} := X_{m-1} \cap T_{m}$ is a submanifold of $\partial T_{m}$,
    triangulated by some faces of $T_{m}$ and their subsimplices.

@@

    On the one hand, if $\Gamma_{m}$ is the entire boundary of $T_{m}$, then it is a topological sphere of dimension $n-1$. 
    Since $\Gamma_{m}$ is a submanifold of the boundary of $X_{m-1}$, which by induction assumption is also a topological sphere of dimension $n-1$,
    we conclude that $\Gamma_{m}$ is the whole boundary of $X_{m}$.
    By basic geometric topology, $X_{m}$ is a topological $n$-sphere, and $T_{m}$ can only be the last simplex in the shelling, $m=M$.
%     then it must also be the boundary of $X_{m-1}$, which is a topological $n$-ball by the induction assumption. 
%     Hence, $\Gamma_{m} = \partial X_{m-1}$ is a topological sphere of dimension $n-1$, 
%     which requires $\partial X_{m-1} \cap \partial T_{m}$ by assumption. 
%     Hence, $X_{m-1} \cup T_{m}$ must be a topological $n$-sphere and thus a manifold without boundary, 
%     which requires $m = M$.
    % DONE{TCF: I don't understand why XM couldn't be a topological n-ball.}
    On the other hand, 
    if $\Gamma_{m}$ is a proper subset of the entire boundary of $T_{m}$, 
    then $X_{m-1} \cup T_{m}$ is still a topological $n$-ball.
\end{proof}

@@

        The vector $a$ is the difference between $\hat h_{G}$ and $\hat h_{K}$ after both are rescaled such that they lie on the affine hyperplane with normal vector $\hat x$. That difference has length at most $\tan(\alpha_{G}) + \tan(\alpha_{K})$.
        We calculate 
        \begin{align*}
            \vecnorm{a}
            = 
            \left\|
            \frac{\vecnorm{ x_{G} }}{\vecnorm{ h_{G} }} \hat h_{{G}} 
            - 
            \frac{\vecnorm{ x_{K} }}{\vecnorm{ h_{K} }} \hat h_K
            \right\|
            &
            \leq 
            \frac{\vecnorm{ x_{G} }}{\vecnorm{ h_{G} }} 
            + 
            \frac{\vecnorm{ x_{K} }}{\vecnorm{ h_{K} }} 
            % \\&
            \leq 
            (\ell+1) 
            \left( 1 + \mu_{T,\ell} \right)
            \sup_{ K \in \patch_{\calT}(S) }
            \aspectratio(K) % \frac{ \diam(K)^{n} }{ n! \vol(K) }
            .
        \end{align*}
        We define 
        \begin{align}
            \nu_{T,\ell} := \sqrt{ 1 + \frac{ (\ell+1)^{2}\left( 1 + \mu_{T,\ell} \right)^{2} \aspectratio(\calT)^{2} }{4} } + \frac{ (\ell+1)\left( 1 + \mu_{T,\ell} \right) \aspectratio(\calT) }{2}
            .
        \end{align}
        In summary, 
        the singular values of the Jacobian of $\Psi$ at almost every $x$ satisfy 
        % using that $\sqrt{s^2+t^2} \leq s + t$ for any $s,t \geq 0$, 
        \begin{align}
            \sigma_{1}(\Psi,x) % \matnorm{ \Jacobian\Psi(x) }
            &\leq  
            (\ell+1)
            \diameterratio(\calT)
            \aspectratio(T)\mu_{T,\ell} % \frac{ \diam(T)^{n} }{ n! \vol(T) }
            \cdot 
            \nu_{T,\ell}
            ,
            \\
            \sigma_{n}(\Psi,x)^{-1} % \matnorm{ \Jacobian\Psi(z)^{-1} }
            &\leq  
            (\ell+1)
            \diameterratio(\calT)
            \aspectratio(\calT) 
            \cdot 
            \nu_{T,\ell}
            ,
            \\
            \sigma_{  2}(\Psi,x) 
            &= 
            \dots
            =
            \sigma_{n-1}(\Psi,x) 
            \leq 
            (\ell+1) \diameterratio(\calT) \aspectratio(T)\mu_{T,\ell}
            ,
            \\
            \sigma_{n-1}(\Psi,x)^{-1} 
            &= 
            \dots
            =
            \sigma_{  2}(\Psi,x)^{-1} 
            \leq 
            (\ell+1) \diameterratio(\calT) \aspectratio(T) 
            ,
            \\
            \det\Jacobian\Psi(x)
            &\leq 
            \left( 
            (\ell+1)
            \diameterratio(\calT)
            \aspectratio(T)\mu_{T,\ell} % \frac{ \diam(T)^{n} }{ n! \vol(T) }
            \right)^{n},
            \\
            \det\Jacobian\Psi^{-1}(x)
            &\leq 
            \left( 
            (\ell+1)
            \diameterratio(\calT)
            \aspectratio(\calT) 
            \right)^{n}
            .
        \end{align}
        We have shown that $\Psi$ and $\Psi^{-1}$ are piecewise Lipschitz with respect to some essentially disjoint closed partitions over their respective domains.
        % It remains to show both are Lipschitz, which we illustrate for $\Psi$, the argument for $\Psi^{-1}$ being analogous. 
        % Rademacher's theorem then shows that both have essentially bounded weak Jacobians. 
        Every function that is piecewise Lipschitz with respect to the triangulation $\calK$ must be Lipschitz over all of $\underlying{\calK}$, since there exists a bi-Lipschitz transformation from $\underlying{\calK}$ onto the unit ball.
        % We use a technique similar to Lemma~\ref{lemma:stars_are_shellable}. 
        % After application of a piecewise linear Lipschitz mapping, 
        % we can assume that $\underlying{\calK}$ is convex without loss of generality. 
        % Any function piecewise Lipschitz with respect to $\calK$ must be Lipschitz over $\underlying{\calK}$. 
        % \mwl{DONE: not enough. Instead, use the homeomorphism onto a ball and estimate its Lipschitz constant.}

@@

        The singular values $\sigma_{\max}(\Theta,G) \geq 1$ and $\sigma_{\min}(\Theta,G) \leq 1$ is monotonically increasing and decreasing, respectively, in $\tan(\beta)$. Hence, they must the extremal maximal singular values of $\Jacobian\Theta_{|G}$.

@@

        Recall that $G \in \calG$ is obtained from $T$ via barycentric subdivisions: 
        first at $z_{S}$, leading to some intermediate $n$-simplex $T'$, and then at $z_{S'}$, leading to $G \subseteq T'$. 
        Let $F_{z} \subseteq G$ be the face opposite to the vertex $z_{S'}$, which is some face of $T'$.
        Now, the height of $F_{z}$ in $G$ is $(n-\ell)$-th of the height of $F_{z}$ in $T'$
        since $S'$ has dimension $n-\ell-1$. 
        The height of the face $F_{z}$ in $T'$ is just the height of its opposing vertex, which lies in $S'$,
        and which equals the height of that vertex in $T$.
%         which must be contained in some face of $T$ containing $S$. 
%         The height vector $h_{z}$ is orthogonal to $F_{z}$. 
%         Since $S'$ has dimension $n-\ell-1$, the vector $(n-\ell) h_{z}$ is the height vector of that face of $T$,
%         pointing to some vertex of $S'$ and standing orthogonally on the face opposite to that vertex. 
        Thus, 
        \begin{align*}
            \frac{ \vecnorm{ b_{z} } }{ \vecnorm{ h_{z} } }
            = 
            \frac{ \vecnorm{ b_{z} } }{ (n-\ell)^{-1} \vecnorm{ (n-\ell) h_{z} } }
            = 
            \frac{ \vecnorm{ z_{S'} } }{ (n-\ell)^{-1} \vecnorm{ (n-\ell) h_{z} } }
            \leq 
            (n-\ell) \aspectratio(T)
%             = 
%             \frac{ \vecnorm{ b_{z} } \vol(F_{z}) }{ (n-\ell)^{-1} n \vol(T) }
%             \leq 
%             \frac{ n-\ell }{n!}
%             \cdot 
%             \frac{ \diam(T)^{n} }{ \vol(T) }
            .
        \end{align*}
        We abbreviate 
        \begin{align}\label{math:upper_bound_for_theta}
            \mu_{T,\ell} 
            := 
            \frac 1 2 \sqrt{ \left( 1 + \rho \right)^{2} + (1+\rho)^{2} (n-\ell)^{2} \aspectratio(T)^{2} } 
            + 
            \frac 1 2 \sqrt{ \left( 1 - \rho \right)^{2} + (1+\rho)^{2} (n-\ell)^{2} \aspectratio(T)^{2} }
            .
        \end{align}
        This establishes bounds on the singular values of the transformation. 
        In summary, the singular values of the Jacobian of $\Theta$ at almost every $x$ satisfy 
        \begin{align}
            \sigma_{1}(\Theta,x) \leq \mu_{T,\ell},
            \quad 
            \sigma_{2}(\Theta,x) = \dots = \sigma_{n-1}(\Theta) = 1,
            \quad 
            \sigma_{n}(\Theta,x)^{-1} = \frac{ \sigma_{1}(\Theta,x) }{ \left| \det \Jacobian\Theta_{|G} \right| } \leq \mu_{T,\ell} / \rho.
        \end{align}
